Machine Learning, Architectural Styles and Property Values

### 2020-11-14 text/manuscript_jrefe.tex
Table updates
manuscript updated

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Machine Learning, Architectural Styles and Property Values},
            pdfauthor={Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson (University of Alabama)},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{soul}
\usepackage{color}
\sethlcolor{yellow}
\usepackage{float}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{multirow}
\doublespacing
\usepackage[backend=bibtex, style=authoryear]{biblatex}
\usepackage[]{biblatex}
\addbibresource{references.bib}

\title{Machine Learning, Architectural Styles and Property Values}
\author{Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson
(University of Alabama)}
\date{14 April 2020}

\begin{document}
\maketitle
\begin{abstract}
This paper introduces an algorithm that collects pictures of individual
buildings from Google Street View. It then trains a deep convolutional
neural network (CNN) to classify residential buildings into
architectural styles, taking into account spatial dependencies of these
styles explicitly. Third, it investigates whether architectural styles
influence house prices. For re-sales, the architectural style is a
significant determinant of transaction prices while no such effect is
found for new buildings. Importantly, this paper explores how to detect
and overcome biases in machine learning methods through a large-scale
comparison of predictions and expert classifications, partially opening
up the `black box' of automatic building classification systems.
\end{abstract}

@@

\footnote{{*Corresponding author: Thies Lindenthal (htl24@cam.ac.uk). Replication files are available from the author's Github repository \href{https://github.com/thies}{https://github.com/thies}. Paul E. Glade and Lukas Heckmann-Umhau are thanked for exceptional research assistance. Carolin Schmidt, Peter Brummund, Mike Langen, Colin Lizieri, Franz Fuerst, Stanimira Milcheva, D'Maris Coffman, Norm Miller, Stephen Malpezzi, John Clapp, Pavel Krivenko and participants at the ASSA/AREUEA 2019 annual meeting and seminars at UCL/Bartlett and the Homer Hoyt Institute offered excellent feedback and critique on earlier versions of this paper.}}

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

If new developments were only pleasing to the eye, so Britain's
\emph{Building Better, Building Beautiful Commission} advising
Government, nimbyism would cease and housing supply could finally reach
the levels demanded by a growing and more affluent population
\autocite{economistlineofbeauty2018}. The inaugural chair of this new
commission boldly suggested that if we built \emph{``as our Georgian and
Victorian forebears built} {[}\ldots{}{]}. \emph{All objections to new
building would slip away in the sheer relief of the public''}
\autocite{scruton2018}. Even Prince Charles, in similar spirit, put
forward ten principles for urban growth and architecture that emphasize
tradition and aesthetics \autocite{princecharles2014}. When governments
and princes occupy themselves with beauty, economists surely may take a
closer look as well. In this paper, we aim to identify the architectural
styles of residential buildings using computer vision techniques and to
empirically search for any transaction price premia associated with
these styles. If aesthetics were as strong a force in the built
environment as claimed, we should indeed find a significant effect on
property prices.

This paper assesses the value of architectural aesthetics head on by
classifying the exterior styles of residential buildings using street
level images. Recent work illustrates how street level imagery and
machine learning classification is an efficient and powerful combination
for measuring previously unobserved characteristics of the urban
environment. \textcite{Naik2017} describe how neighborhood demographics
may impact the physical appearance of neighborhoods.
\textcite{Gebru2017} use classified vehicle make and model information
to predict income, race, education, and voting patterns at the precinct
level. \textcite{Glaeser2016} predict income in New York City.
\textcite{Naik2016AER} create a neighborhood safety based Streetscore
which is shown to be highly correlated with neighborhood population
density and household income. \textcite{Nadai2016} find that greenery
and street facing windows contribute to a positive appearance of safety
while \textcite{Liu2017} evaluate the quality and upkeep of the built
environment along Beijing's streets.

\textcite{GlaeserKincaidNaik2018} push the level of observation from the
block, street, or street-section level to the individual
\emph{building level}. Utilizing images of buildings' exteriors
collected from Google Street
View\footnote{\href{https://www.google.co.uk/maps}{https://www.google.co.uk/maps}},
and to a lesser degree interior images from Zillow, they find that looks
matter, at least in Boston: A one standard deviation improvement of a
building's exterior is associated with an additional USD 70,000 in home
value. Intuitively, the link between good looks and value is
bi-directional: The appearance of buildings that went through
foreclosure deteriorated significantly
\autocite{GlaeserKincaidNaik2018}.

Zooming in at individual buildings significantly increases the benefits
of using mass collected street level imagery in economic research:
property characteristics previously deemed ``unobservable'' can be
directly observed in an accurate, objective, automatic and therefore
cost effective way. Deriving additional variables from unconventional
data sources like 3D airborne laser scanning (or in our case Google
Street View) is fundamental since this provides ``essential determinants
influencing real estate prices {[}which{]} are constantly missing and
are not accessible in official and mass appraiser databases''
\autocite{Helbich2013}.

In this paper we suggest a method to collect a large number of images of
individual UK buildings from Google Street View, classify the depicted
buildings using deep convolutional neural networks, combine the derived
information with sales price data and, ultimately, estimate marginal
prices for the estimated building characteristics. Our work demonstrates
that utilizing pretrained convolutional neural networks to detect
complex characteristics such as housing appearances with relatively high
accuracy comes at low computational costs and is feasible with only
modestly-sized training data. Additionally, we illustrate how machine
classifications may be sensitive to photo quality by comparing machine
classifications to those made by human experts.

\hypertarget{building-styles-and-property-values}{%
\subsection{Building styles and property
values}\label{building-styles-and-property-values}}

The aesthetics of different architectural styles make for an easy
cocktail party conversation topic -- and they may also influence market
prices. Using architectural assessments by human experts,
\textcite{Buitelaar2017} find a sizable premium of 5\% for new buildings
in the Netherlands that refer to traditional styles and a staggering
15\% premium for new buildings that closely follow traditional shapes,
facade composition and details. Their study carefully disentangles the
architectural style from other unobserved characteristics such as
building quality, differences in location, or year of construction.
These controls are crucial, as earlier work has established that age and
style variables tend to be highly correlated. \textcite{Coulson2008},
for instance, suggest a non-parametric estimator and establish a
U-shaped age function and distinct price discounts for postwar and
contemporary styles (vis-à-vis more historic styles).
\textcite{Francke2017a} investigate the depreciation of residential real
estate in the Netherlands and decompose land versus structure values
singling out the effect of ``physical deterioration, functional
obsolescence and and vintage effects''. They find that buildings from
the 1930s carry a strong price premium.

A large scale assessment of buildings' exteriors will allow for an
analysis of the externalities of architecture. Buildings hardly ever
stand in isolation and \textcite{Ahlfeldt2012} investigate the influence
a building's architecture exerts on its surroundings. They observe a
positive price effect for residential buildings in the direct proximity
of iconic homes by Frank Lloyd Wright in Oak Park, Illinois. A
building's exterior does not need to be an architectural masterpiece to
co-determine the value of other houses close by. Homogeneity of building
shapes within street segments does influence property values. A
similarly shaped neighboring building is value enhancing while proximity
to a wildly different neighboring shape, everything else remaining
equal, is detrimental to property values \autocite{Lindenthal2017b}.

Unfortunately, the traditional approaches chosen by
e.g.~\textcite{Buitelaar2017} or \textcite{Ahlfeldt2012} do not scale
well. Each observation has to be classified into architectural styles by
a human expert, which is time-consuming and costly, imposing an upper
limit on the number of observations and the level of detail captured for
each observation in any given sample. Google Street View offers a
solution to this nexus as it captures the images of almost all buildings
in many cities around the world at high level of accuracy and detail.
The challenge we address in this paper is to extract and utilize
building level information from this ubiquitous sensor, using freely
available deep learning techniques.

\hypertarget{ml-classification-of-building-styles}{%
\section{ML classification of building
styles}\label{ml-classification-of-building-styles}}

@@

The first challenge when trying to collect building images in the UK is
fundamental: How can one identify the building of interest correctly?
Simply looking up an address on Google Maps too often leads to imprecise
`street views' and not the building-level portraits needed. Fig.
\ref{fig:84Vin} presents a typical result from an address level search,
showing a broad ensemble of buildings instead of singling out the
building of interest (in this case, the partly captured terraced house
at the very left margin).

@@

Previous work sourcing imagery from Google Street View has mostly
focused on neighborhood or precinct characteristics where exact spatial
assignment of objects is not required and street-level images suffice.
For major US cities, the accuracy of building image search results is
higher, which should help studies such as
\textcite{GlaeserKincaidNaik2018}.

Self-evidently, the images we use for classifying architectural style
must focus on specific buildings of interest. For the UK, and many other
countries, the Google Street View API returns the coordinates of the
nearest camera snapshot for a given location but fails to provide an
accurate orientation and zoom-level of the camera needed to capture the
front of the building. To estimate the best possible view, we combine
Street View metadata queries and Ordnance Survey maps to find the
optimal camera orientation to center images on the front door of the
building of interest.

@@

Specifically, we first look up the nearest Google Street View panorama
point (green dot in Figure \ref{fig:fan}) based on the centroid (red
dot) coordinates of a given building obtained from Ordnance Survey maps.
We then perform a viewshed analysis and identify which exterior walls
are visible from the panorama point, ignoring any wall segments where
the direct line of sight from the panorama point is obstructed by other
buildings. We can then estimate the camera bearing (green line) and zoom
factor, based on the fan of the lines of sight (in
blue).\footnote{Code necessary to replicate this analysis is available from the authors' GitHub.com code repositories.}

Obstructed lines of sight due to greenery, fences, garden walls or large
vehicles cannot be detected from the Ordnance Survey maps. We therefore
use a first stage image classification procedure to identify if the
building image is obstructed and we proceed the second-closest location
if the line of sight is obstructed.

By combining land registry maps with the snapshot location, we are able
to automatically build a set of building frontage images for
approximately 48,000 properties in Cambridge (UK). The Ordnance Survey
maps do not provide any use type classifications so we collect images on
all buildings that have a ground plate between 40 and 250 \(m^2\). This
method, however, could easily be adapted to photograph buildings in
locations for which high-resolution building level maps are not
available by using LIDAR based building outlines instead, which would be
in the spirit of \textcite{Glaeser2016}.

\hypertarget{from-pixels-to-computer-vision}{%
\subsubsection{From pixels to computer
vision}\label{from-pixels-to-computer-vision}}

We download all pictures at the highest resolution offered by Google's
Street View
API\footnote{\href{https://developers.google.com/maps/documentation/streetview/intro}{https://developers.google.com/maps/documentation/streetview/intro}}
(640x640 color pixels). We then obtain 2048-dimensional feature vectors
for each picture, using the \emph{Inception-v3} deep convolutional
neural network \autocite{Szegedy2015}. \emph{Inception-v3} has been
trained for the ImageNet Large Scale Visual Recognition Challenge
(ILSVRC)\footnote{\url{http://image-net.org/challenges/LSVRC/}}, which
evaluates image classification and object detection algorithms for a
wide range of objects. The pretrained classifications would allow us to
identify pets, vehicles or people on the pictures -- assessing
architectural style, however, is beyond the canned classifiers'
capabilities.

\textcite{GlaeserKincaidNaik2018} rely on an different ILSVRC
competitor, \emph{Resnet-101} \autocite{He2016}. They reduce the
extracted feature vectors to lower dimensionality (1024 to 100
dimensions) based on principal component analysis (PCA). We follow a
different strategy and double the dimensionality by including the
feature vector of the closest building in the sample, which, in most
cases, is the direct neighbor. The feature vectors of neighbors have
been collected in exactly the same way as those for other buildings in
the sample. Doubling up allows us to model spatial dependencies in
building styles, similar to spatially correlated land cover
classifications in \textcite{Ghimire2010a}.

\hypertarget{image-classification}{%
\subsection{Image classification}\label{image-classification}}

Cambridge's houses can be classified into seven broad styles:
\footnote{We are grateful to colleagues from the Architecture Department at the University of Cambridge to provide these general style descriptions.}

>>>>

\documentclass[]{article}

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}

\usepackage[a4paper, margin=1in]{geometry}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Machine Learning, Architectural Styles and Property Values},
            pdfauthor={Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson (University of Alabama)},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{soul}
\usepackage{color}
\sethlcolor{yellow}
\usepackage{float}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{multirow}
\doublespacing
\usepackage[backend=bibtex, style=authoryear]{biblatex}
\usepackage[]{biblatex}
\addbibresource{bibliography.bib}

\begin{document}

\title{Machine Learning, Architectural Styles and Property Values}
\author{Thies Lindenthal* (University of Cambridge) \& Erik B. Johnson
(University of Alabama)}
\date{7 November 2020}

\maketitle
\begin{abstract}
This paper couples a traditional hedonic model with architectural style classifications from human experts and machine learning (ML) enabled classifiers to estimate sales price premia over architectural styles, both at the building and the neighborhood-level. We find statistically and economically significant price differences for houses from distinct architectural styles across an array of specifications and modeling assumptions. Comparisons between classifications from ML models and human experts illustrate the conditions under which ML classifiers may perform at least as reliable as human experts in mass appraisal models. Hedonic estimates illustrate that the impact of architectural style on price is attenuated by properties with less well-defined styles and we find no evidence for differential price effects of \emph{Revival} or \emph{Contemporary} architecture for new construction.
\end{abstract}

@@

\footnote{{*Corresponding author: Thies Lindenthal (htl24@cam.ac.uk). Replication files are available from the author's Github repository \href{https://github.com/thies}{https://github.com/thies}. Paul E. Glade and Lukas Heckmann-Umhau are thanked for their exceptional research assistance. Carolin Schmidt, Peter Brummund, Mike Langen, Colin Lizieri, Franz Fuerst, Stanimira Milcheva, D'Maris Coffman, Norm Miller, Stephen Malpezzi, John Clapp, Pavel Krivenko and participants at the ASSA/AREUEA 2019 annual meeting and seminars at UCL/Bartlett and the Homer Hoyt Institute offered excellent feedback and critique on earlier versions of this paper.}}

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}
In this paper, we introduce a new data collection technique that allows us to rigorously test the notion that households have strong preferences over architectural styles. Our results indicate that 1) there is evidence for sales price premia associated with houses from a variety of architectural styles, 2) hedonic estimates from high confidence machine learning (ML) enabled architectural image classifications are similar to human expert estimates, and 3) any style premia are found for existing structures only and not for new buildings, indicating that debates about strong demand for historicizing aesthetics may be unfounded.

Understanding household preferences for architectural styles is of increasing interest to both policymakers and researchers. The policy perspective is best illustrated by Britain's \emph{Building Better, Building Beautiful Commission}, which advises the government on design choices for homes and neighborhoods. It lists as one of its primary aims as \emph{``To make the planning system work in support of better design and style, not against it.''} \autocite{economistlineofbeauty2018} and boldly claims that matching the style of new housing to aggregate neighborhood preferences should overcome otherwise prevailing objections of incumbent households to new construction.

Researchers have recently begun to quantify and define what makes for \emph{better design and style}. \textcite{Buitelaar2017} find evidence on the link between preferences over architectural styles and housing prices using architectural assessments by human experts, and estimate a sizable premium of 5\% for new buildings in the Netherlands that refer to traditional styles and a staggering 15\% premium for new buildings that closely follow traditional shapes, facade composition, and details. \textcite{Coulson2008} suggest a non-parametric estimator and establish a U-shaped age function and distinct price discounts for postwar and contemporary styles (vis-à-vis more historic styles). \textcite{Francke2017a} investigate the depreciation of residential real estate in The Netherlands and decompose land versus structure values singling out the effect of physical deterioration, functional obsolescence, and vintage effects. They find that buildings from the 1930s carry a strong price premium. These results support the hypothesis that preferences for house vintages (which coincide with architectural styles) may affect housing prices.

Clearly, building vintage, quality, location, and architectural styles are highly correlated which implies that hedonic price estimates for architectural styles cannot simply be understood as marginal prices for aesthetics. We only interpret our empirical findings as `pure' architectural preferences in few cases where differences in all other dimensions can be accounted for, e.g. new units at similar locations that have been built to similar standards and that have not depreciated yet. For markets with more contemporary housing stocks, differences in functional obsolescence would be a major concern when age and style variables are correlated. In our case, over 93\% of the housing stock had been produced prior to 1980 and 70\% before 1940,  so concerns on contemporaneous depreciation of new structures may be alleviated as the vast majority of homes has already depreciated. Nevertheless, systematic differences in locations and unobservable quality characteristics and modernization standards cannot be ruled out.

\textcite{Buitelaar2017} indicate that any premium for an architectural style must stem from either differences in construction prices or from supply
constraints, as new construction potentially does not capture the demand
for traditional styles. In markets where home builders are free to decide on the architectural style of the homes they supply, we would not expect any price effects beyond differences in construction costs. If there was, for instance, a premium for more traditional styles, developers would continue to build more of these until all demand is met. Our prior is therefore \emph{not} to find distinct price effects of aesthetics for newly supplied homes.

Of course, residential buildings rarely stand in isolation and \textcite{Ahlfeldt2012} investigate the influence a building's architecture exerts on its surroundings. They observe a positive price effect for residential buildings in the direct proximity of iconic homes by Frank Lloyd Wright in Oak Park, Illinois. A building's exterior does not need to be an architectural masterpiece to co-determine the value of other houses close by. A similarly shaped neighboring building is value-enhancing while proximity to a wildly different neighboring shape, everything else remaining equal, may be detrimental to property values \autocite{Lindenthal2017b}.

In traditional approaches chosen by e.g.~\textcite{Buitelaar2017} or \textcite{Ahlfeldt2012}, each observation is classified into an architectural style by a human expert, which is time-consuming, costly and is not feasible for large sample sizes. As \textcite{Helbich2013} state, the benefit of using new methods to observe the built environment is that ``essential determinants influencing real estate prices {[}which{]} are constantly missing and are not accessible in official and mass appraiser databases''. This is certainly the case in our context, as understanding the impact of architectural style on housing prices is made difficult by the paucity of sales or assessment data that includes architectural style as a characteristic. In addition, an expert's views may not fully reflect the market's perceptions of relevant styles -- architectural classifications and economically relevant segments might differ.

Combining automated Google Street View photo data collection with deep learning/ML image classification offers a promising way forward to large and comprehensive data sets of architectural style. \textcite{Naik2017} describe how neighborhood demographics may impact the physical appearance of neighborhoods. \textcite{Gebru2017} use classified vehicle make and model information to predict income, race, education, and voting patterns at the precinct level. \textcite{Glaeser2016} predict income in New York City. \textcite{Naik2016AER} create a neighborhood safety based Streetscore which is shown to be highly correlated with neighborhood population
density and household income. \textcite{Nadai2016} find that greenery
and street-facing windows contribute to a positive appearance of safety
while \textcite{Liu2017} evaluate the quality and upkeep of the built
environment along Beijing's streets.

In contrast to the block, street, or street-section level classifications used in the majority of previous studies that use automated image classification, \textcite{GlaeserKincaidNaik2018} push the level of observation to the individual \emph{building level}. Utilizing images of buildings' exteriors collected from Google Street View\footnote{\href{https://www.google.co.uk/maps}{https://www.google.co.uk/maps}}, and to a lesser degree interior images from Zillow, they find that looks matter, at least in Boston: A one standard deviation improvement of a building's exterior is associated with an additional USD 70,000 in home value. Intuitively, the link between good looks and value is
bi-directional: The appearance of buildings that went through foreclosure deteriorated significantly \autocite{GlaeserKincaidNaik2018}.

We follow the framework used by \textcite{GlaeserKincaidNaik2018} and focus on individual buildings as our unit of observation. This focus unlocks one of the main benefits of using mass collected street-level imagery in economic research: property characteristics previously deemed ``unobservable'' can be directly observed in an increasingly accurate and objective manner and may be implemented in a more efficient, accessible, and cost-effective way.

In addition to reducing the cost of data collection, the ability for low-cost classification of a large sample of structures in an urban area enables researchers to detect not only the architectural style of the building but also to characterize the style of other buildings in the vicinity. This allows us to examine the interaction between the style of a building and that of its neighbors.

To analyze the accuracy of our method of detecting architectural style, we compare our algorithmic results with an extensive data set of architectural style classifications compiled by human experts. The relatively large and costly comparison group provides insights into ML-based classification accuracy and the robustness of hedonic estimates to human and ML classifications. In addition, the measurable uncertainty in the ML classifications may also be a useful source of variation to exploit for style identification.  For example, it may be the case that the impact of architectural style on sales price for existing buildings may be attenuated for buildings that are more difficult to classify. To our knowledge, this step is not present in the existing literature and provides additional insight into the efficacy of the technique for application in other domains.

 We find evidence that for existing buildings, the architectural style has an impact on the sales price and that the estimates are very similar for both human expert and high confidence ML classified images. The price effects of human-expert defined architectural styles are greater in images where there is agreement with the ML classifications. The presence of architectural style in our hedonic specifications reduces the RMSE of our models. The preferred architectural style for resales appears to be \emph{Revival} with \emph{Interwar}, \emph{Postwar}, and \emph{Early Victorian} commanding the smallest premium. Immediately proximate neighbor styles also have an impact on sales price, with Contemporary neighborhoods clearly preferred to Georgian, Early Victorian, and Postwar. Importantly for policymakers, we do not identify a premium for either Revival or Contemporary styles for new construction.

We begin by describing the traditional hedonic data used in the paper. Second, we describe the methodology for the automated architectural style classification. Third, we compare the results of the automated predictions with expert classifications. Next, we present an array of hedonic regressions using both architect and ML-based classification and discuss results. Finally, we conclude with a discussion of the policy implications of the estimates and the feasibility of ML-based classification of architectural style.

\hypertarget{Data}{%
\section{Data}\label{data}
}

Residential real estate transactions in England and Wales are
collected and published by the Land Registry
\autocite{LandRegistry2016a}. Their records include the date of transaction, the price paid, street address, a classification of the
property type (flat, detached, semi-detached, or terraced house), the
estate type (freehold or leasehold), and an indicator for newly built
properties. We select transactions from Cambridge, England, which were recorded
between January 1995 and October 2018, excluding any leaseholds, apartments
and properties classified as type ``other'', and sales with prices below
£50,000 or in excess of £2,000,000. Notably, the Land Registry data lack relevant variables such as year of construction, home and lot size, or similar quality indicators. We, therefore, augment the transaction data with core hedonic variables from other sources and to control for a variety of building-specific effects that may be correlated with both sales price and architectural style.

The Ordnance Survey (OS)\footnote{\url{https://www.ordnancesurvey.co.uk/}} provides high-resolution maps that show the two-dimensional footprint of buildings in the UK. Using these maps, we calculate each building's floor plate (in \(m^2\)) and estimate its volume by combining the building outlines with digital elevation models from the \textcite{EnvironmentAgency2015}, as suggested by \textcite{Lindenthal2017}. Additionally, we measure the distance from each residential property to the city center, proxied by Great St.~Mary's Church. The \textcite{ONS2017} subdivides Cambridge into 69 unique \emph{Lower Super Output Areas} (LSOA) and we rely on these boundaries when constructing neighborhood dummy variables. An LSOA typically has 1,000--3,000 residents and 400--1,200 households of comparable economic and socio-demographic characteristics \autocite{ONS2017b}.

\hypertarget{ml-classification-of-building-styles}{%
\subsection{Architectural Style Data}\label{ml-classification-of-building-styles}}

Our main variable of interest, architectural style, follows the classifications commonly used by realtors, home-buyers, and architects. Styles are distinctive in Cambridge: Table \ref{tab:styles_rubric} provides a description of the construction eras and the distinctive motifs for each style as defined by members of the Architecture Department at the University of Cambridge.

\begin{center}
  -- Insert Table \ref{tab:styles_rubric} about here --
\end{center}

We aim to collect images of all residential buildings in Cambridge from Street View. Approximately 50,000 buildings found on the OS maps have footprints between 30 and 500 $m^2$. For 85\% of these candidates, we can download street-level images on Google Street View. However, not all of them turn out to be residential buildings. A sample of 25,000 of these images is categorized by final year architecture students into the seven styles.

 We can match 23,768 residential properties sold in Cambridge between 1995 and 2019 to images from Google Street View. We implicitly assume that houses have not changed their architectural styles between the time of sale and the time Google collected the images. This is a reasonable approach since conversions of just the exterior of buildings rarely occur. In case of full redevelopments, we exclude any sales for that address pre-dating the redevelopment, since we cannot rule out a change of style at redevelopment.

 Importantly, 15,511 sales can also be matched to the sample of expert's classifications. Our hedonic models will be estimated on these observations for which we have predicted and ground truth classifications. This allows us to assess the prediction accuracy and robustness of the regression results to the classification approach. When assessing the architectural style of surrounding buildings, we use the full set of all images collected. Being able to cover almost the universe of all buildings in a market is a clear advantage of an automated approach.

In Figure \ref{fig:vint_examples}, we present a set of representative images for each of the 7 architectural styles. These images come from our automated collection procedure and are a visual representation corresponding to the expert descriptions in Table \ref{tab:styles_rubric}. Image information here is very useful as it provides a visual mapping from the style characteristics to actual identification. One component which is evident in the images but not necessarily in the descriptions is the within style homogeneity of window grouping, relative pane size, and counts. For example, Georgian windows contain 12 equally sized panes in 4 rows and 3 columns, Early Victorian windows consist of 4 equally sized panes in 2 rows and 2 columns, Interwar windows contain an upper smaller pane and larger lower pane. Interestingly, the Revival style demonstrates heterogeneous window patterns that borrow from the variety of classic architectural styles. We revisit the impact of image quality and window visibility on prediction accuracy below.

@@

To provide the best information to detect style for both the architect and ML classifiers, images should be focused on the structure under examination, have a clear field of view, show minimal picture overlap with neighboring properties, and minimize potentially confounding issues such as  vehicle type.

Unfortunately, in the UK, Google's Street View API regularly fails to identify the building at a given address. Instead, the camera location will be in close proximity to the front of the building but aimed straight ahead and down the center of the road. Fig. \ref{fig:84Vin} (a) presents a typical result from an address level API. Here, the Google Street View API algorithm fails to accurately capture the front of the building and the property is not even contained in the photo. For major US cities, the accuracy of Google's search results is higher and the Street View camera will pan towards the center of the property parcel. This makes the photo collection of studies such in U.S urban areas potentially much more straightforward although the API may result in images that are too 'zoomed out' resulting in images that may contain obstructed views, vehicles, and neighboring properties.

@@

Our automated image collection algorithm utilizes a sophisticated process of viewshed analysis based on Ordnance Survey building maps and ML methods to capture images using the best camera locations, zoom, and angles while accounting for potential obstructions from buildings and trees or vehicles. Thus, for each property, our goal is to automate the collection of the best possible Street View image for each property in our sample.

More specifically, we first use OS maps to identify each building's outline (see Figure \ref{fig:fan}). Second, we build a database of all possible Google Street View camera locations using Metadata queries of addresses in the Ordnance Survey.  These metadata queries return the latitude, longitude, and dates of the most recent panorama photos collected by Google Street View. Next, we use the OS data to assign each property to the nearest Street View photo location in the metadata database and conduct a viewshed analysis on surrounding building outlines to assign pan and zoom parameters to find an informative view. The algorithm allows us to estimate the camera bearing (green line) and zoom factor, based on the fan of the lines of sight (in blue). The line of sight criteria identifies which exterior walls are visible from the nearest Street View point and aims away from any wall segments where the direct line of sight is obstructed by other buildings. A visual example of this algorithm is shown in  Figure \ref{fig:fan}.\footnote{Code necessary to replicate the full image capture including metadata capture and viewshed based camera aiming is available from \url{https://github.com/thies/paper-cambridge-vintages}}. For each property, we then collect the image at the highest resolution offered by Google's Street View API\footnote{\href{https://developers.google.com/maps/documentation/streetview/intro}{https://developers.google.com/maps/documentation/streetview/intro}} (640x640 color pixels). We test the robustness of our hedonic models to a variety of image quality concerns in the results section of the paper.

@@

Figure \ref{fig:84Vin} (b) shows how the suggested approach can improve the quality of the harvested images. The new image is focused appropriately on the relevant property, it contains minimal neighbor information. and the view is not overly obstructed by other structures, greenery, or vehicles. As a result, the image contains the relevant identifying architectural style characteristics such as a clearly identified roofline, front door, window panes, and bay window.

\hypertarget{image-classification}{%
\section{Methodology: Style classification}\label{image-classification}}

With the images collected, we next focus on training an ML learning model to classify the architectural style observed in each of the collected Street View images. The first step in building our ML model is to construct a training data set. Our training data set consists of images from 25,000 randomly sample properties classified into one of seven different architectural styles by final year students from the Architecture Department at the University of Cambridge\footnote{The full set of classified images are available for download at the authors' websites.} A sample of 25,000 is much more training data than is actually needed. By applying transfer learning to an existing image recognition model, we require less than 250 observations per category to reach saturated training accuracy levels. We use this ``surplus'' training data to perform out of sample comparisons between our ML predictions and human expert classifications. This large set of expert classified data also allows us to perform robustness, and falsification checks on the hedonic regressions for the 15,369 overlapping observations where we have transaction data with both human and ML-based style classifications.

\hypertarget{transfer-learning}{%
\subsection{Training the model}\label{transfer-learning}}

We simplify the calibration of our architectural style prediction model by utilizing transfer learning. Transfer learning freezes the parameters from a pre-trained image recognition model, removes the final steps in the model, and then trains a new model on the vector of outputs from the truncated pre-trained model to produce new classifications\footnote{For examples, see \url{https://www.tensorflow.org/tutorials/images/transfer_learning}}. Transfer learning methods enable the use of smaller training data sets and impose a much lower computational burden than traditional deep learning models. 
For our purposes, we use the \emph{Inception-v3} deep convolutional
neural network \autocite{Szegedy2015}. \emph{Inception-v3} has been
trained for the ImageNet Large Scale Visual Recognition Challenge
(ILSVRC)\footnote{\url{http://image-net.org/challenges/LSVRC/}}, which
evaluates image classification and object detection algorithms for a
wide range of objects. The pre-trained classifications would allow us to
identify pets, vehicles, or people in the pictures -- assessing
architectural style, however, is beyond the canned classifiers'
capabilities. To allow for the detection of new objects, transfer learning techniques strip off the last steps of the model. The output of the original deep learning model is essentially just a deeply recursive dimension reduction engine that outputs a 2048-dimensional feature vector for each picture. This feature vector describes the outline shapes, locations, and colors that were important for the ImageNet classifications.

\textcite{GlaeserKincaidNaik2018} rely on a different ILSVRC competitor, \emph{Resnet-101} \autocite{He2016}. They reduce the extracted feature vectors to lower dimensionality (1024 to 100 dimensions) based on principal component analysis (PCA). In our analysis, we follow a different strategy and actually double the dimensionality (4096) of our feature vectors by including the feature vector of the closest building in the sample, which, in most cases, is the direct neighbor. The feature vectors of neighbors have been collected in exactly the same way as those for other buildings in the sample. Doubling up allows us to model spatial dependencies in building styles, similar to spatially correlated land cover classifications in \textcite{Ghimire2010a}.
